---
title: "Tools for Attribute VSI CUSUM Charts"
author: "Hanwool Kim"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tools for Attribute VSI CUSUM Charts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The package attrCUSUM has been developed to aid in 
the statistical design of attribute VSI CUSUM chart. 
It currently provides tools for monitoring of mean increase when the data are
(zero inflated) Poisson/binomial distribution. These tools can be easily applied
to other count processes and more complicated CUSUM control chart. 
We leave these issues as our perpetual work.

This vignette briefly illustrates some examples 
for usage of functions in the package. 

## CUSUM statistic
Note that 
the considered CUSUM statistic for monitoring of mean increase is defined as follows:

${C}_{0} = {c}_{0},$

${C}_{t} = \max \left( 0, {C}_{t-1}\right) + X_{t} - k, ~~ t \in \mathbf{N}$,

wherein $X_{t}$ is a count process of interest, 
$k$ is the reference value and $C_{0}$ is the initial state.

## Getting ANSSs, ATSs and other informations
First of all, the package should be attached.
```{r attach}
library(attrCUSUM)
```

Suppose that $X_{i}$ is distributed Poisson with mean $\lambda$. 
The function `getAve_poisson` provides useful information 
on design of upper one-sided CUSUM control chart when the data are Poisson. 

Assume that we are interested in 
the average number of sample to signal (ANSS) an out-of-condition for 
CUSUM(k=5,h=8) with initial state 0 when the data are Poisson with mean 4 where 
k and h are the reference value and 
the (upper) control limit of CUSUM control chart respectively. 
Then the ANSS of CUSUM(k=5, h=8) chart can be obtained as follows:
```{r ANSS_pois_1}
lambda <-  4  # mean of Poisson distribution
k <- 5  # reference value 
h <- 8  # control limit
c.zero <- 0  # initial state
res <- getAve_pois(lambda = lambda, refv = k, contl = h, c.zero = c.zero)
res$ANSS
```

Next, Assume that we want to get the corresponding longer sampling time when 
the predetermined shorter sampling interval is 0.1 and the warning limit is -1.
The results can be obtained as follows:

```{r ATS_pois_1}
lambda <-  4  # mean of Poisson distribution
k <- 5  # reference value 
h <- 8  # control limit
c.zero <- 0  # initial state
warnl <- -1  # warning limit
ds <- 0.1  # shorter sampling interval
res1 <- getAve_pois(lambda = lambda, refv = k, contl = h, 
                   c.zero = c.zero, warnl = warnl, ds = ds)
dl <- res1$dl  # longer sampling interval
dl
res2 <- getAve_pois(lambda = lambda, refv = k, contl = h, 
                   c.zero = c.zero, warnl = warnl, ds = ds, dl = dl)
res2$ANSS
res2$ATS
```

The following is an example of ATS profiles. 
Evaluate `example(package = "attrCUSUM", "getAve_pois")` for more details.
In plot, $\delta$ denotes mean shift in units of standard deviation.
```{r ATS_profiles_pois_1, echo=FALSE, fig.align='center', fig.width=7, fig.height=5}
lambda <-  4  # mean of Poisson distribution
k <- 5  # reference value 
h <- 8  # control limit
c.zero <- 0  # initial state
warnl <- -1  # warning limit
std.lambda <- sqrt(lambda)
dss <- c(1, 0.75, 0.5, 0.25, 0.1)  # shorter sampling intervals
deltas <- seq(0, 2.0, 0.1)  # shift  
for (ds in dss) {
  dl <- getAve_pois(lambda = lambda, refv = k, contl = h,
                    ds = ds, warnl = warnl)$dl
  ds.text <- as.character(ds)
  ats.text <- paste("ats", ds.text, sep = "_")
  eval.text.1 <- paste(ats.text, " <- numeric(length(deltas))", sep = "")
  eval.text.2 <- paste(ats.text, "[j] <- res$ATS", sep = "")
  eval(parse(text = eval.text.1))
  for (j in seq_along(deltas)) {  
    res <- getAve_pois(lambda = lambda + (std.lambda * deltas[j]), 
                       refv = k,contl = h, ds = ds, dl = dl, warnl = warnl)
    eval(parse(text = eval.text.2))
  }
}
plot(x = deltas, y = log10(ats_0.1), ylim = c(-0.5, log10(ats_1)[1]),
     pch = 0, type = "o", main = "VSI Poisson CUSUM(5,8)", cex.main = 1.1,
     sub = paste("w=", warnl, sep = ""), cex.sub = 1.1,
     xlab = expression(delta), cex.lab = 1.1,
     ylab = expression(paste(log[10], ATS)))
lines(x = deltas, y = log10(ats_0.25), pch = 1, type = "o")
lines(x = deltas, y = log10(ats_0.5), pch = 6, type = "o")
lines(x = deltas, y = log10(ats_0.75), pch = 7, type = "o")
lines(x = deltas, y = log10(ats_1), pch = 16, type = "o")
legend(x = 1.4, y = 2.25, legend = paste("ds=", dss, sep = ""),
       pch = c(16, 7, 6, 1, 0), cex = 0.9, bty = "n")

```

Note that the function `getAve_pois` provides other useful information 
related to Markov chain approach. 
Refer to `help(getAve)` for more details.
```{r names_getAve_pois}
names(res2)
```

## Getting suitable control limit
As was before, Assume that $X_{i}$ is distributed Poisson with mean $\lambda$, 
also we want to get suitable control limit when 
the data are Poisson with mean 4 for that 
the reference value is 5, the initial state is 0 and 
the desired ANSS is near 200. 
Then, the suitable control limit can be obtained as follows:
```{r Contl_pois_1}
lambda <-  4  # mean of Poisson distribution
k <- 5  # reference value 
c.zero <- 0  # initial state
anss.target <- 200L  # desired anss
getContl_pois(lambda = lambda, anss.target = anss.target, 
              refv = k, c.zero = c.zero)

```

## Other count process
Although we illustrated only the case that the data are Poisson,
the package `attrCUSUM` also provides other count processes.

* For (zero-inflated) binomial VSI CUSUM chart, 
refer to `help(getAve_binomial)` and `help(getContl_binomial)`.
* For (zero-inflated) Poisson VSI CUSUM chart, 
refer to `help(getAve_Poisson)` and `help(getContl_Poisson)`.
* For user-specified VSI CUSUM chart, 
refer to `help(getAve)` and `help(getContl)`.

## Additional issues
* The core computational algorithms are implemented using 'Rcpp' and 'RcppArmadillo'.
* For user-friendly arguments checking, the package 'ArgumentCheck' is used.
* 'attrCUSUM' provides functions for density, distribution, quantile and
    random generation for the zero inflated Poisson distribution. Refer to `help(zipois)`
* 'attrCUSUM' provides functions for density, distribution, quantile and
    random generation for the zero inflated binomial distribution. Refer to `help(zibinom)`





